### 2013-10-12 research/MCR report/main.tex
uploaded 12-10-2013

\section{Introduction}
The search for alternative energy sources which could be categorised under the ”green”label has become important area of research in the modern word. Solar, wind power and wave power are some of the examples of these sources. Recently,a new branch of research has been developing to extract energy from flow induced vibrations. It has been hypothesized that this technique may work efficiently in areas where regular turbines cannot.
A simple structure that is susceptible to flow-induced vibrations that are suitable for energy extraction are slender structures,such as cylinders, elastically mounted perpendicular to a fluid stream. With regards to a slender body two common types of   flow induced vibrations are Vortex Induced Vibrations (VIV) and aeroelastic galloping. Significant research has been carried out by Bernitsas and his team on extracting useful energy from VIV. Some of their significant work includes investigating the influence  of physical parameters such as mass ratio (the ratio of the mass of the cylinder and the displaced fluid), Reynolds number, mechanical properties(\cite{Raghavan2010a} ,\cite{Lee2011b}) and location (effect of the bottom boundary) (\cite{Raghavan2009}). However,the possibility of extracting energy using aeroelastic galloping has not been thoroughly investigated. Some theoretical work was carried out by (\cite{Barrero-Gil2010a}). Utilizing galloping may be a more viable method to harness energy from flow induced vibrations as it is not bounded by a "lock-in" range of reduced velocities(ratio between the freestream velocity and the product of the natural frequency of the system and the characteristic length).Therefore it is preferable to investigate further the possibility of harnessing energy from flow induced vibrations using aeroelastic galloping.

@@

In order to carry out an optimisation study on the fluid dynamics portion of the system, it is essential to study underlying fluid mechanics for galloping. It has been already established by \cite{Parkinson1964} that the instantaneous lift is the input of the system. Therefore it is essential to know how exactly this lift occur. According to \cite{Parkinson1989},\hilight{Luo 94} the lift force is crated by the secondary flow of the two shear layers of either side of the body which creates an unbalanced pressure distribution on the afterbody. If a square cross section is taken as an example as the cross section have a transverse $\dot{y}$, the relative velocity causes an instantaneous angle of attack which makes one shear layer lies closer to the body than the other. Therefore the shear layer near the body creates a pressure imbalance by creating a higher suction on the wall of the body on the adjacent side than the suction caused by the shear layer on the opposite side on its adjacent side of the body. The lift force become its maximum when the shear layer closer to the boy re attaches at the trailing edge. As the angle of attack is increased beyond that point, The separated shear layer in the opposite side of the body (which was initially not close to the body) shrinks therefore the pressure difference decreases which results a decrease in the lift force. This explanation was validated by \cite{Luo1994} by experimental flow visualisation.

>>>>

\section{Introduction}
The search for alternative energy sources which could be categorised under the ”green” label has become important area of research in the modern word. Solar, wind power and wave power are some of the examples of these sources. Recently,a new branch of research has been developing to extract energy from flow induced vibrations. It has been hypothesized that this technique may work efficiently in areas where regular turbines cannot.
A simple structure that is susceptible to flow-induced vibrations that are suitable for energy extraction are slender structures,such as cylinders, elastically mounted perpendicular to a fluid stream. With regards to a slender body two common types of   flow induced vibrations are Vortex Induced Vibrations (VIV) and aeroelastic galloping. Significant research has been carried out by Bernitsas and his team on extracting useful energy from VIV. Some of their significant work includes investigating the influence  of physical parameters such as mass ratio (the ratio of the mass of the cylinder and the displaced fluid), Reynolds number, mechanical properties(\cite{Raghavan2010a} ,\cite{Lee2011b}) and location (effect of the bottom boundary) (\cite{Raghavan2009}). However,the possibility of extracting energy using aeroelastic galloping has not been thoroughly investigated. Some theoretical work was carried out by (\cite{Barrero-Gil2010a}). Utilizing galloping may be a more viable method to harness energy from flow induced vibrations as it is not bounded by a "lock-in" range of reduced velocities(ratio between the freestream velocity and the product of the natural frequency of the system and the characteristic length).Therefore it is preferable to investigate further the possibility of harnessing energy from flow induced vibrations using aeroelastic galloping.

@@

In order to carry out an optimisation study on the fluid dynamics portion of the system, it is essential to study underlying fluid mechanics for galloping. It has been already established by \cite{Parkinson1964} that the instantaneous lift is the input of the system. Therefore it is essential to know how exactly this lift occur. According to \cite{Parkinson1989},\hilight{Luo 94} the lift force is crated by the secondary flow of the two shear layers of either side of the body which creates an unbalanced pressure distribution on the afterbody. If a square cross section is taken as an example as the cross section have a transverse $\dot{y}$, the relative velocity causes an instantaneous angle of attack which makes one shear layer lies closer to the body than the other. Therefore the shear layer near the body creates a pressure imbalance by creating a higher suction on the wall of the body on the adjacent side than the suction caused by the shear layer on the opposite side on its adjacent side of the body. The lift force become its maximum when the shear layer closer to the body re attaches at the trailing edge. As the angle of attack is increased beyond that point, The separated shear layer in the opposite side of the body (which was initially not close to the body) shrinks therefore the pressure difference decreases which results a decrease in the lift force. This explanation was validated by \cite{Luo1994} by experimental flow visualisation.

### 2013-10-12 research/paper/figure/m_star_collapsed.tex
uploaded 12-10-2013

\begin{figure}
\setlength{\unitlength}{\textwidth}

  \begin{picture}(1,0.35)(0,0.75)

  \put(0.15,0.76){\includegraphics[width=0.7\unitlength]{../FnP/gnuplot/mean_power_collapsed_mstar.eps}}

 	\put(0.11,0.95){\large $\frac{P_{m}}{\rho \mathcal{A}U^3 }$}

 	 	\put(0.45,0.75){ $c\rho\mathcal{A}U$}

  \end{picture}

@@

>>>>

\begin{figure}

\setlength{\unitlength}{\textwidth}
\fbox{
  \begin{picture}(1,0.32)

       \put(0.07,0.03){\includegraphics[width=0.3\unitlength]{../FnP/gnuplot/mean_power_collapsed_mstar.eps}}
       \put(0.36,0.03){\includegraphics[width=0.3\unitlength]{../FnP/gnuplot/mean_power_collapsed_noshed_mstar.eps}}
       \put(0.65,0.03){\includegraphics[width=0.3\unitlength]{../FnP/gnuplot/fsi_power.eps}}

        \put(0.03,0.17){\large $\frac{P_{m}}{\rho \mathcal{A}U^3}$}

        \put(0.2,0.0){$\frac{c}{\rho\mathcal{A}U}$} 	
        \put(0.49,0.0){$\frac{c}{\rho\mathcal{A}U}$}
        \put(0.78,0.0){$\frac{c}{\rho\mathcal{A}U}$}

        \put(0.133,0.21){\small(a)}
        \put(0.42,0.21){\small(b)}
        \put(0.715,0.21){\small(c)}

  \end{picture}
  }

@@

Mean power as a function of damping factor at Re

### 2013-10-12 research/paper/results/results-2.3.tex
uploaded 12-10-2013

 The data presented in Fig.\ref{fig:uncollapsed_data} for various damping ratios, $\zeta$, can be collapsed into a single line for a for a particular force characteristic curve (i.e $C_y$ vs $\theta$ curve). These collapsed curves were  obtained for the velocity amplitude  and power by plotting as functions of as a function of  the non dimensionalised  damping constant $c\rho\mathcal{A}U$ 
(Fig \ref{fig:collpased_data} (a),(b),(c) and (d)).  This further emphasizes that the galloping system is not frequency dependent. It is possible to obtain a different similar power output at different values of \ustar when the damping constant, $c\rho\mathbb{A}U$, is kept fixed. An example of this case, as shown in Fig.\ref{fig:time_hostory_velocity_same_power}, clearly show that this is a result of similar velocity amplitudes between cases if one were to disregard the high frequencies due to shedding. As mentioned earlier, it is the transverse velocity that determines the energy provided by the fluid forcing and the mechanical damping.

@@

  At region 3 ($U^*= 400$) `$c$' is low in comparison with region 1 and 2 which leads to a low mean power output. Fig.\ref{fig:power_time_histories} (c) shows that $P_t$ becomes negative over some portion of the cycle. This is because $\theta$  passes the point where both $\theta$ and $C_y$ (therefore $F_y$) are positive. The high velocity amplitude leads to the equivalent incident angle $\theta$, in this case to exceed the range where $C_y$ is positive (i.e. $0<\theta<6^0$ in Fig.\ref{cy ploynomial} (a)). In this portion of the cycle the hydrodynamic force actually opposes the direction of travel and power is transferred from the structure to the fluid during those times. From and energy perspective, the mechanical damping is not sufficient to remover the energy transferred from the fluid to the structure during other times of the cycle because $c\rho\mathcal{A}U$ is substantially low. Therefore this excess energy is transferred back to the fluid as depicted by the negative region of $P_d$ in Fig.\ref{fig:power_time_histories} (c)

@@

In this paper, the power transfer of a square body under aero elastic galloping is analysed by using quasi-steady state model and numerical methods. The system is not frequency dependent but more dependent on the velocity amplitude of the system. It could be proven from the collapsed plots in terms of non-dimensionalised (using flow parameters) damping constant. By anaylizing key regions of the power vs \ustar curve it could be concluded that in order to obtain an optimum power output, the damping constant ($c\rho\mathcal{A}U$) should be high, but not too high enough to hinder the induced angle of attack reaching the region of higher forcing. The reduction in power at low mass ratios was due to the fact that the instantaneous $P_t$ and $P_d$ are not the same. Therefore the mechanical system requires adequate inertia to store a portion of power transferred from the fluid to the body to sustain galloping.

>>>>

 The data presented in Fig.\ref{fig:uncollapsed_data} for various damping ratios, $\zeta$, can be collapsed into a single line for a for a particular force characteristic curve (i.e $C_y$ vs $\theta$ curve). These collapsed curves were  obtained for the velocity amplitude  and power by plotting as functions of as a function of  the non dimensionalised  damping constant $\frac{c}{\rho\mathcal{A}U}$ 
(Fig \ref{fig:collpased_data} (a),(b),(c) and (d)).  This further emphasizes that the galloping system is not frequency dependent. It is possible to obtain a different similar power output at different values of \ustar when the damping constant, $c\rho\mathbb{A}U$, is kept fixed. An example of this case, as shown in Fig.\ref{fig:time_hostory_velocity_same_power}, clearly show that this is a result of similar velocity amplitudes between cases if one were to disregard the high frequencies due to shedding. As mentioned earlier, it is the transverse velocity that determines the energy provided by the fluid forcing and the mechanical damping.

@@

  At region 3 ($U^*= 400$) `$c$' is low in comparison with region 1 and 2 which leads to a low mean power output. Fig.\ref{fig:power_time_histories} (c) shows that $P_t$ becomes negative over some portion of the cycle. This is because $\theta$  passes the point where both $\theta$ and $C_y$ (therefore $F_y$) are positive. The high velocity amplitude leads to the equivalent incident angle $\theta$, in this case to exceed the range where $C_y$ is positive (i.e. $0<\theta<6^0$ in Fig.\ref{cy ploynomial} (a)). In this portion of the cycle the hydrodynamic force actually opposes the direction of travel and power is transferred from the structure to the fluid during those times. From and energy perspective, the mechanical damping is not sufficient to remover the energy transferred from the fluid to the structure during other times of the cycle because $\frac{c}{\rho\mathcal{A}U}$ is substantially low. Therefore this excess energy is transferred back to the fluid as depicted by the negative region of $P_d$ in Fig.\ref{fig:power_time_histories} (c)

@@

In this paper, the power transfer of a square body under aero elastic galloping is analysed by solving the quasi-steady state model equations through numerical integration. The power output of the system is not dependent on \ustar or natural frequency of the system, but dependent on the transverse velocity amplitude of the system. This is shown by the collapsed plots in terms of non-dimensionalised damping constant (using flow parameters). By analysing key regions of the power vs \ustar curve it could be concluded that in order to obtain an optimum power output, the damping constant ($\frac{c}{\rho\mathcal{A}U}$) should be high, but not excessive until it  to hinders the galloping from reaching induced angles of attack where the forcing is significant.
